---
title: "Lab 4"
author: "Kaitlin Smith"
date: "11/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(corrplot)
library(purrr)
```

## TL;DR

Quick overview: 
1. I didn't find any outliers

2. I created the feature with the average NDAI score that we talked about in lab.
Right now, the code takes about an hour to run for the 3 images, so I'd recommend downloading the data I uploaded 
(if anyone could improve it, that would be great!)

3. Data:

each image is stored in a dataframe: img1, img2, and img3

I added the following features:

1. label_final: this changed all of the "0"- unlabeled pixels to "1" - cloud pixels

2. Avg_NDAI: average NDAI score for the [(x-4,y-4),(x+4),(y+4)] box around each pixel (this was a little tricky to calcultae, because 
the images aren't perfectly square. If a 9x9 square could not be created for a pixel, the largest area around that pixel wihtin the constraint was found and the average was taken for whatever number of pixels was in the area.)

3. Sub_feat: I tried to replicate the NDAI metric by finding (DF*CF - AF*AN) / (DF*CF + AF*AN). It looks like this 
didn't work great by the correlation matrices 

4. Mult_Rads: I multiplied DF, CF, BF, AF, AN together for each pixel. 
The rational behind this was that each radiation angle appears to be 
hightly correlated with each other

5. Mult_AvgNDAI_SD: I multiplied the Avg_NDAI by SD. It doesn't look
like this improved the AvgNDAI metric 




## EDA

```{r}
feature_names <- c("x", "y", "label", "NDAI", "SD","CORR", "DF", "CF", "BF", "AF", "AN")
img1 <- read.table("data/image_data/image1.txt")
colnames(img1) <- feature_names
img2 <- read.table("data/image_data/image2.txt")
colnames(img2) <- feature_names
img3 <- read.table("data/image_data/image3.txt")
colnames(img3) <- feature_names
```

## Basic EDA/ Check if we need to do any Data Cleaning

Quick check for outliers:

```{r}
generate_boxplot <- function(data){
  boxplot(data)
}

invisible(apply(img1[,4:11], MARGIN = 2, generate_boxplot)) # use invisible to suppress boxplot printed output

invisible(apply(img2[,4:11], MARGIN = 2, generate_boxplot))

invisible(apply(img3[,4:11], MARGIN = 2, generate_boxplot)) 
```
It appears that there aren't any clear outliers/ missing data. This won't be 
in the final report, but I kept this in if anyone was curious of if you 
wanted to see the distributions of all of the features. 


```{r fid.width = 20}

heatmap_ggplot <- function(index, img){
  return(ggplot(img, aes(x = x,y = y, fill = img[,index])) + geom_tile() + labs(title = colnames(img)[index]))
}

maps <- lapply(3:11, heatmap_ggplot, img1)

ggarrange(plotlist = maps)
```

```{r}
maps2 <- lapply(3:11, heatmap_ggplot, img2)

ggarrange(plotlist = maps2)
```

```{r}
maps3 <- apply(3:11], MARGIN = 2, heatmap_ggplot, img3)

ggarrange(plotlist = maps3)
```
## Rational for Converting Unlabeled pixels to "Cloud" Labels 

In the plots bellow, it appears that the distribution of the cloud labels 
is contained in the labels that were unlabeled. For this reason, 
we will consider the unlabeled pixels as cloudy pixels. 

```{r}
# TODO: CHANGE COLORS

ndai_label1 <- ggplot(img1, aes(x = NDAI, fill = factor(label))) + geom_density(alpha = 0.4)+ labs(title = "Image1") + scale_fill_manual(labels=c("Not Cloud", "Unlabeled", "Cloud"), values = c("blue", "red", "green"))

ndai_label2 <- ggplot(img2, aes(x = NDAI, fill = factor(label))) + geom_density(alpha = 0.4) + labs(title = "Image2") + scale_fill_manual(labels=c("Not Cloud", "Unlabeled", "Cloud"), values = c("blue", "red", "green"))

ndai_label3 <- ggplot(img3, aes(x = NDAI, fill = factor(label))) + geom_density(alpha = 0.4) + labs(title = "Image3") +scale_fill_manual(labels=c("Not Cloud", "Unlabeled", "Cloud"), values = c("blue", "red", "green"))

ggarrange(ndai_label1, ndai_label2, ndai_label3)
```
Below, I'll add a "label_final" feature that converts all of the 
unlabeled pixels to cloud labels 

```{r}
img1$label_final <- replace(img1$label, x == 0, 1)
img2$label_final <- replace(img2$label, x == 0, 1)
img3$label_final <- replace(img3$label, x == 0, 1)

```



## Correlation Matrices, Two Ways

```{r}

M <- cor(img1)

corrplot(M, method="number")

M2 <- cor(img2)

corrplot(M2, method="number")

M3 <- cor(img3)

corrplot(M3, method="number")
```

You can see in the above correlation matrices, that it appears that NDAI
is the feature that is most correlated with the expert label. 
For image1 and image2, it appears that the angles of radiance also have a negative correlation. However, image3 does not have this relationship. 
with the labels. In addition, it appears that NDAI and SD have a positive correlation.

## Feature Engineering 

First, here's a summary of the features that have already been engineered 
for us. Note that for each radiance angle measurement at each pixel, there are 64 radiation measurements.

1. CORR:

This feature is an average of the average linear combination 
between the Af/An and Bf/An radiance angle pairs. Clouds produce
low correlation measures, unless the clouds are in motion. Higher correlation
indicates no clouds or a low altitude cloud. 

2. SD:

This feature is computed by finding the standard deviation within groups
of the An radiance angle. This feature can help identify smooth
surfaces, which are typical of clouds. However, if the ground 
also had a smooth surface, this feature is not as informative.

3. NDAI:

This feature is computed by finding the difference in the average Df and 
average An radiance angle for each pixel, divided by the sum of these two numbers.


### Average NDAI 

From the correlation matrix, it appeared that NDAI was the feature 
that was most correlated with the labels. Intuitively, the spacial aspect 
of this data is important. We expect "cloud" pixels and "not cloud" pixels 
to be clustered together. Or, if a label is "cloud", we would expect 
the other pixels around it to also be labeled as "cloud". Thus, we'll
construct a feature whose values for a given pixel is the average of the 
NDAI values for the 8 pixels around the pixel. If a pixel is on the edge of 
the image, the minimum value of $(x, mod(x + 4, max(x val)))$ is taken. 


```{r}
# # converted the dataframe to matrices for speed-up
# data <- as.matrix(img1[c("x", "y", "NDAI")])
# get_average_matrix <- function(index){
# 
#   this_y <- data[index, 2]
#   this_x <- data[index, 1]
#   # get all of the columns with x and y values in the range we want. This works even if the area we find is not square (because the data isn't #   square). Then find the mean of the NDAI column for this region
# 
#   avg <- mean((data[data[,1] >= this_x - 4 & data[,1] <= this_x + 4 & data[,2] >= this_y - 4 & data[,2] <= this_y + 4, ])[,3]) 
#   
#   return(avg)
# }
# 
# avg_NDAI_img1 <- lapply(1:nrow(img1), get_average_matrix)

```
```{r}
# img2_matrix <- as.matrix(img2[c("x", "y", "NDAI")])
# get_average_matrix_img2 <- function(index){
# 
#   this_y <- img2_matrix[index, 2]
#   this_x <- img2_matrix[index, 1]
# 
#   avg <- mean((img2_matrix[img2_matrix[,1] >= this_x - 4 & img2_matrix[,1] <= this_x + 4 & img2_matrix[,2] >= this_y - 4 & img2_matrix[,2] <= this_y + 4, ])[,3])
#   
#   return(avg)
# }
# 
# img3_matrix <- as.matrix(img3[c("x", "y", "NDAI")])
# get_average_matrix_img3 <- function(index){
# 
#   this_y <- img3_matrix[index, 2]
#   this_x <- img3_matrix[index, 1]
# 
#   avg <- mean((img3_matrix[img3_matrix[,1] >= this_x - 4 & img3_matrix[,1] <= this_x + 4 & img3_matrix[,2] >= this_y - 4 & img3_matrix[,2] <= this_y + 4, ])[,3])
#   
#   return(avg)
# }
# 
# avg_NDAI_img2 <- lapply(1:nrow(img2), get_average_matrix_img2)
# 
# avg_NDAI_img3 <- lapply(1:nrow(img3), get_average_matrix_img3)

```



```{r}
# img1$Avg_NDAI <- unlist(avg_NDAI_img1)
# 
# img2$Avg_NDAI <- unlist(avg_NDAI_img2)
# 
# img3$Avg_NDAI <- unlist(avg_NDAI_img3)
```

### Other Features

```{r}

modify_rad <- function(DF, CF, AF, AN){
  return((DF*CF - AN*AF)/(DF*CF + AN*AF))
}

mult_rad <- function(DF, CF, BF, AF, AN){
  return(DF * CF * AF * AN)
}

mult_ndai_sd <- function(ndai, sd){
  return(ndai * sd)
}

# new features for image 1

img1 <- img1 %>% mutate(Sub_Feat = unlist(pmap(img1[c("DF", "CF", "AF", "AN")], function(DF, CF, AF, AN) modify_rad(DF = DF, CF = CF, AF = AF, AN = AN))))

img1 <- img1 %>% mutate(Mult_Rads = unlist(pmap(img1[c("DF", "CF", "BF", "AF", "AN")], function(DF, CF, BF, AF, AN) mult_rad(DF = DF, CF = CF, BF = BF, AF = AF, AN = AN))))

img1 <- img1 %>% mutate(Mult_AvgNDAi_SD = unlist(pmap(img1[c("Avg_NDAI", "SD")], function(Avg_NDAI, SD) mult_ndai_sd(ndai = Avg_NDAI, sd = SD))))

# repeat for image 2
img2 <- img2 %>% mutate(Sub_Feat = unlist(pmap(img2[c("DF", "CF", "AF", "AN")], function(DF, CF, AF, AN) modify_rad(DF = DF, CF = CF, AF = AF, AN = AN))))

img2 <- img2 %>% mutate(Mult_Rads = unlist(pmap(img2[c("DF", "CF", "BF", "AF", "AN")], function(DF, CF, BF, AF, AN) mult_rad(DF = DF, CF = CF, BF = BF, AF = AF, AN = AN))))

img2 <- img2 %>% mutate(Mult_AvgNDAi_SD = unlist(pmap(img2[c("Avg_NDAI", "SD")], function(Avg_NDAI, SD) mult_ndai_sd(ndai = Avg_NDAI, sd = SD))))

# repeat for image 3
img3 <- img3 %>% mutate(Sub_Feat = unlist(pmap(img3[c("DF", "CF", "AF", "AN")], function(DF, CF, AF, AN) modify_rad(DF = DF, CF = CF, AF = AF, AN = AN))))

img3 <- img3 %>% mutate(Mult_Rads = unlist(pmap(img3[c("DF", "CF", "BF", "AF", "AN")], function(DF, CF, BF, AF, AN) mult_rad(DF = DF, CF = CF, BF = BF, AF = AF, AN = AN))))

img3 <- img3 %>% mutate(Mult_AvgNDAi_SD = unlist(pmap(img3[c("Avg_NDAI", "SD")], function(Avg_NDAI, SD) mult_ndai_sd(ndai = Avg_NDAI, sd = SD))))

```



```{r fig.width = 15}

M <- cor(img1)

corrplot(M, method="number")

M2 <- cor(img2)

corrplot(M2, method="number")

M3 <- cor(img3)

corrplot(M3, method="number")
```

```{r}
# save data
write.csv(img1, "data/img1_features.csv", row.names = FALSE)
write.csv(img2, "data/img2_features.csv", row.names = FALSE)
write.csv(img3, "data/img3_features.csv", row.names = FALSE)
```

